---
title: "Running an experiment"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Running an experiment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(treat.sim)
library(simmer)
```

## Creating an experiment for the model

In `treat.sim` all experimental parameters are held in a `list`. You can use the convenience function `create_experiment` to create a pre-populated list with all the correct parameter names and default values.

> use help('create_experiment') to see the full list of parameters available. 

```{r}
default_experiment <- create_experiment()
print(default_experiment$n_triage_bays)
print(default_experiment$triage_mean)
```

The above example creates a default experiment. To vary parameters simply set the corresponding argument in the `create_experiment` function.

```{r}
extra_triage <- create_experiment(n_triage_bays=2)
print(extra_triage$n_triage_bays)

extra_faster_triage <- create_experiment(n_triage_bays=2,
                                         triage_mean=2.75)

print(extra_faster_triage$n_triage_bays)
print(extra_faster_triage$triage_mean)
```

## Run the model and view a log of events

To run the model pass a simmer environment and experiment list to the `single_run` function. To see a log of events set the environments log level to 1.

> To reduce the printed log we se the `run_length` of the model to 100 time units (minutes). The default run length of the model is 1020 minutes.

```{r}
default_experiment <- create_experiment()

env <- simmer("TreatSimmer", log_level=1)
env <- single_run(env, default_experiment, run_length=100.0)
```

To run the model and suppress the log exclude the `log_level`

```{r}
default_experiment <- create_experiment()

env <- simmer("TreatSimmer")
env <- single_run(env, default_experiment, run_length=100.0)
```
