---
title: "Model code"
author: "Thomas Monks, Alison Harper"
format: html
editor: visual
---

## 1. Imports

> Note: we are calculating KPIs using our own code here, but you can also use `simmer.plot`. Help with install of `simmer.plot` (`igraph` installation is the actual issue) on Linux based systems: <https://r.igraph.org/articles/installation-troubleshooting#cannot-compile-igraph-from-sources-on-linux>
>
> > If you make use of conda environments (via Anaconda/mini-conda/mini-forge/mamba), **remember** to `conda deactivate` before installation.

```{r}
library(simmer)
library(simmer.bricks)
suppressMessages(library(simmer.plot))
library(tibble)
library(ggplot2)
suppressMessages(library(RCurl))
suppressMessages(library(Rlab))
suppressMessages(library(dplyr))
suppressMessages(library(tidyr))

options(dplyr.summarise.inform = FALSE)
```

## 2. Default values and constants

### 2.1 Distribution parameters

```{r}
#' Mean and Variance of the underlying Normal Distribution
#' 
#' @description
#' `normal_moments_from_lognormal` calculates the mu and sigma
#' of the normal distribution underlying a lognormal
#' mean and standard 
#'
#' @details
#' `rlnorm` from `stats` is designed to sample from the lognormal distribution. 
#' The parameters is expects are moments of the underlying normal distribution
#' Using sample mean and standard deviation this function calculates 
#' the mu and sigma of the normal distribution. source: https://blogs.sas.com/content/iml/2014/06/04/simulate-lognormal-data-with-specified-mean-and-variance.html
#' 
#' @param mean A number. Sample mean.
#' @param stdev A number. Sample standard deviation
#' @returns A list 
normal_moments_from_lognormal <- function(mean, std){
  phi <- sqrt(std^2 + mean^2)
  mu <- log(mean**2/phi)
  sigma <- sqrt(log(phi^2/mean^2))
  return(list("mu" = mu, "sigma" = sigma))
}
```

```{r}
# sign-in/triage parameters
DEFAULT_TRIAGE_MEAN <- 3.0

# registration parameters (lognormal distribution)
DEFAULT_REG_PARAMS <- normal_moments_from_lognormal(5.0, sqrt(2.0))

# examination parameters
DEFAULT_EXAM_PARAMS = list(mean=16.0, var=3.0)

# trauma/stabilisation
DEFAULT_TRAUMA_MEAN <- 90.0

# Trauma treatment (lognormal distribution)
DEFAULT_TRAUMA_TREATMENT_PARAMS <- normal_moments_from_lognormal(30.0, sqrt(4.0))

# Non trauma treatment (lognormal distribution)
DEFAULT_NON_TRAUMA_TREATMENT_PARAMS <- normal_moments_from_lognormal(13.3, sqrt(2.0))

# prob patient requires treatment given trauma
DEFAULT_NON_TRAUMA_TREAT_P <- 0.60

# proportion of patients triaged as trauma
DEFAULT_PROB_TRAUMA <- 0.12
```

### 2.2 Time dependent arrival rate data

The data for arrival rates varies between clinic opening at 6am and closure at 12am.

```{r}
# data are held in the Github repo and loaded from there.
NSPP_PATH = 'https://raw.githubusercontent.com/TomMonks/open-science-for-sim/main/src/notebooks/01_foss_sim/data/ed_arrivals.csv'

csv_data <- getURL(NSPP_PATH)
df <- read.csv(text=csv_data)

# lock in order of time of day for bar chart display
df$period <- factor(df$period, levels = df$period)

ggplot(data=df, aes(x=period, y=arrival_rate)) +
  geom_bar(stat="identity", fill="steelblue") + 
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust=1)) +
  xlab("Hour of day") + 
  ylab("Mean arrivals (patients/hr)")

```

### 2.3 Resource Counts

Integer count variables representing the number of resources at each activity in the process

```{r}
DEFAULT_N_TRIAGE <- 1
DEFAULT_N_REG <- 1
DEFAULT_N_EXAM <- 3

# stabilisation rooms
DEFAULT_N_TRAUMA <- 10000

# Non-trauma cubicles
DEFAULT_NON_TRAUMA_CUBICLES <- 1

# trauma pathway cubicles
DEFAULT_TRAUMA_CUBICLES <- 1
```

### 2.4 Simulation model run settings

```{r}
# Random seed - this will be investigated for CRN
SEED <- 42

# default results collection period
DEFAULT_RESULTS_COLLECTION_PERIOD <- 60 * 19

# number of replications.
DEFAULT_N_REPS <- 5

# Show the a trace of simulated events
# 1 = show, 0 = do not show.
LOG_LEVEL <- 1
```

## 3. Functions

Load and format data

```{r}
load_arrival_data <- function(path=NSPP_PATH){
  csv_data <- getURL(NSPP_PATH)
  df <- read.csv(text=csv_data)
  
  # arrivals per minute...
  df$arrival_rate2 <- df$arrival_rate/60.0
  
  # create 60 minute increments for period
  df$period = seq(0, (nrow(df)-1)*60, by=60)
  return(df)
}
```

```{r}
#' Sample a patient type
#' 
#' @description
#' `sample_arrival_type` samples if a patient type is trauma or non-trauma
#' with a given probability.
#'
#' @details
#' The function uses the Bernouli distribution (Rlab) to sample
#' if a patient is Trauma or Non-Trauma.  The return values are 
#' 1 = Trauma, 2 = Non-trauma.
#' @param p A number: the probability a patient has trauma on arrival
sample_arrival_type <- function(p, n=1){
  ifelse(rbern(n, prob = DEFAULT_PROB_TRAUMA) == 1, 1, 2)
}
```

```{r}
#' Sample a if a non-trauma patient requires treatment
#' 
#' @description
#' `sample_nt_trauma_treatment` samples if a non-trauma patient
#' requires cubicle treatment
#'
#' @details
#' The function uses the Bernouli distribution (Rlab) to sample
#' if a patient is requires treatment or not.  The return values are 
#' 1 = Treatment, 0 = No treatment
#' @param p A number: The probability the patient requires treatment
sample_nt_trauma_treatment <- function(p){
  ifelse(rbern(1, prob = p) == 1, 1, 0)
}
```

Sampling from a non-stationary poisson process using thinning

```{r}
nspp_thinning <- function(simulation_time, data, debug=FALSE){
  
  # calc time interval: assumes intervals are of equal length
  interval <- data$period[2] - data$period[1]
  
  # maximum arrival rate (smallest time between arrivals)
  lambda_max <- max(data$arrival_rate2)

  while(TRUE){
    # get time bucket (row of dataframe to use)
    t <- floor(simulation_time / interval) %% nrow(data) + 1
    lambda_t <- data$arrival_rate2[t]
    
    # set to a large number so that at least 1 sample is taken
    u <- Inf
    rejects <- -1
    
    # running total of time until next arrival
    inter_arrival_time <- 0.0
    
    # reject proportionate to lambda_t / lambda_max
    ratio <- lambda_t / lambda_max
    while(u >= ratio){
      rejects <- rejects + 1
      # sample using max arrival rate
      inter_arrival_time <- inter_arrival_time + rexp(1, lambda_max)
      u <- runif(1, 0.0, 1.0)
    }
    
    if(debug){
      print({paste("Time:", simulation_time, 
                   " Rejections:", rejects, 
                   " t:", t, 
                   " lambda_t:", lambda_t, 
                   " IAT:", inter_arrival_time)})
    }
      
    return(inter_arrival_time)
  }
}
```

## 4. Model parameterisation

The model is setup to be created from a set of functions that return trajectories. Each function accepts a list that contains all parameters to configure the simulation model. Here we create the list and pre-populate it using default values.

```{r}
create_experiment <- function(n_triage_bays=DEFAULT_N_TRIAGE,
                              n_reg_clerks=DEFAULT_N_REG,
                              n_exam_rooms=DEFAULT_N_EXAM,
                              n_trauma_rooms=DEFAULT_N_TRAUMA,
                              n_non_trauma_cubicles=DEFAULT_NON_TRAUMA_CUBICLES,
                              n_trauma_cubicles=DEFAULT_TRAUMA_CUBICLES,
                              triage_mean=DEFAULT_TRIAGE_MEAN,
                              stabilisation_mean=DEFAULT_TRAUMA_MEAN,
                              trauma_treat_params=DEFAULT_TRAUMA_TREATMENT_PARAMS,
                              reg_params=DEFAULT_REG_PARAMS,
                              exam_params=DEFAULT_EXAM_PARAMS,
                              prob_non_trauma_treat=DEFAULT_NON_TRAUMA_TREAT_P,
                              nontrauma_treat_params=DEFAULT_NON_TRAUMA_TREATMENT_PARAMS,
                              prob_trauma=DEFAULT_PROB_TRAUMA,
                              arrival_data_path=NSPP_PATH,
                              log_level=LOG_LEVEL) {
  
  # load arrival data
  arrival_data <- load_arrival_data(path=arrival_data_path)
  
  # create list of parameters
  experiment <- list(n_triage_bays=n_triage_bays,
                    n_reg_clerks=n_reg_clerks,
                    n_exam_rooms=n_exam_rooms,
                    n_trauma_rooms=n_trauma_rooms,
                    n_non_trauma_cubicles=n_non_trauma_cubicles,
                    n_trauma_cubicles=n_trauma_cubicles,
                    triage_mean=triage_mean,
                    stabilisation_mean=stabilisation_mean,
                    trauma_treat_params=trauma_treat_params,
                    reg_params=reg_params,
                    exam_params=exam_params,
                    prob_non_trauma_treat=prob_non_trauma_treat,
                    nontrauma_treat_params=nontrauma_treat_params,
                    prob_trauma=prob_trauma,
                    arrival_data=arrival_data,
                    log_level=log_level)
  
  return(experiment)
}     
```

## 5. Patient Trajectories

The DES package `simmer` uses the concept of a `trajectory` to model a process for a particular patient type. In the urgent care centre example trajectories allow us to model separate trauma and non-trauma processes. Note that different trajectories can share common resources.

The `simmer` terminology for using resources and engaging in activities is easy to read:

-   `seize` - queue and take a resource when it is available.

-   `timeout` - a process delay (e.g. treatment or diagnostics)

-   `release` - release a resource.

`simmer` also provides a way to set an attribute of the `trajectory` using `set_attribute`. This is useful for storing timing information to display in a log: for example when a patient begins waiting for a resource (access via `now(env)`).

> **Important notes:**
>
> 1.  The function `log_` is used in combination with `function()` `paste` to provide a dynamic simulation trace to the R console.
> 2.  Sampling code should look as follows:
>
> ``` r
> timeout(task = function() rexp(1, 3.0)) %>%
> ```
>
> > The keyword `function()` must be included for dynamic sampling for each patient. **Omitting** `function()` means that it is evaluated **once** at the time the `trajectory` is created.

### 5.1. Trauma Patients

> We wrap the trajectory in a function called `create_trauma_pathway`. This allows us to pass an argument `exp` that can parameterise the trajectory for use in a discrete experiment.

```{r}

create_trauma_pathway <- function(exp){
    
    trauma_pathway <- trajectory(name="trauma_pathway") %>%
      set_attribute("patient_type", 1) %>%
      # log patient arrival
      log_(function() {paste("**Trauma arrival")}, level=1) %>% 
      
      # triage 
      set_attribute("start_triage_wait", function() {now(exp$env)}) %>%
      visit("triage_bay", function() rexp(1, 1/exp$triage_mean)) %>%
      log_(function() {paste("(T) Triage wait time:",
                       now(exp$env) - get_attribute(exp$env, "start_triage_wait"))},
           level=1) %>%
      
      # request trauma room for stabilization
      set_attribute("start_trauma_room_wait", function() {now(exp$env)}) %>%
      visit("trauma_room", function() rexp(1, 1/exp$stabilisation_mean)) %>%
      log_(function() {paste("(T) Trauma room wait time:",
                       now(exp$env) - get_attribute(exp$env, "start_trauma_room_wait"))},
           level=1) %>%
      
      # request treatment cubicle
      set_attribute("start_trauma_treat_wait", function() {now(exp$env)}) %>%
      visit("trauma_treat_cubicle", function() rlnorm(1, exp$trauma_treat_params$mu,
                                                      exp$trauma_treat_params$sigma)) %>%
      log_(function() {paste("********************(T) Trauma treatment cubicle wait time:",
                       now(exp$env) - get_attribute(exp$env, "start_trauma_treat_wait"))},
           level=1) %>% 
    
      # store the total time in system 
      set_attribute("total_time", 
                    function() {now(exp$env) - get_attribute(exp$env, "start_triage_wait")})
    
  return(trauma_pathway)
}
```

### 5.2 Non-trauma patients

```{r}

create_nt_cubicle_treatment <- function(exp){

  nt_cubicle_treatment <- trajectory() %>% 
    log_(function() {paste("NT patient requirement treatment")},
         level=1) %>% 
    seize(resource="nontrauma_treat_cubicle", amount=1) %>% 
    
    timeout(task = function() rlnorm(1, exp$nontrauma_treat_params$mu,                                                     exp$nontrauma_treat_params$sigma)) %>%
    release(resource = "nontrauma_treat_cubicle", amount = 1) %>% 
    log_(function() {paste("NT treatment complete")},
     level=1) %>% 
  return(nt_cubicle_treatment)
}
```

```{r}

create_non_trauma_pathway <- function(exp){
  # log messages
  ARRIVAL_MSG = "**Non-Trauma arrival**"
  TRIAGE_MSG = "(NT) Triage wait time:"
  REG_MSG = "Reg wait time:"
  EXAM_MSG = "Exam wait time:"
  EXIT_MSG = "NT Total time in system:"
  
  # optional trajectory for proportion of patients that requirement treatment
  nt_cubicle_treatment <- create_nt_cubicle_treatment(exp)
  
  non_trauma_pathway <- trajectory(name="non_trauma_pathway") %>%
    set_attribute("patient_type", 2) %>%
    # log non_trauma arrival
    log_(function() {paste(ARRIVAL_MSG)}, level=1) %>% 
    
    # store start of waiting time for log calculations
    set_attribute("start_triage_wait", function() {now(exp$env)}) %>%
    # queue and use triage bay
    visit("triage_bay", function() rexp(1, 1/exp$triage_mean)) %>%
    log_(function() {paste(TRIAGE_MSG, now(exp$env) - get_attribute(exp$env, "start_triage_wait"))},
         level=1) %>%
    
    # queue and use registration clerk
    set_attribute("start_reg_wait", function() {now(exp$env)}) %>%
    visit("registration_clerk", function() rlnorm(1, exp$reg_params$mu, 
                                                  exp$reg_params$sigma)) %>%
    log_(function() {paste(REG_MSG, now(exp$env) - get_attribute(exp$env, "start_reg_wait"))},
         level=1) %>%
    
    # queue and use examination room
    set_attribute("start_exam_wait", function() {now(exp$env)}) %>%
    visit("examination_room",  function() rnorm(1, exp$exam_params$mean, 
                                                sqrt(exp$exam_params$var))) %>%
    log_(function() {paste(EXAM_MSG, now(exp$env) - get_attribute(exp$env, "start_exam_wait"))},
         level=1) %>%
    
    # a Proportion of patients require treatment in a cubicle
    branch (
      function() sample_nt_trauma_treatment(exp$prob_non_trauma_treat), continue=T,
      nt_cubicle_treatment
    ) %>% 
    log_(function() {paste(EXIT_MSG, now(exp$env) - get_attribute(exp$env, "start_triage_wait"))},
         level=1) %>% 
    # store the total time in system 
    set_attribute("total_time", 
                    function() {now(exp$env) - get_attribute(exp$env, "start_triage_wait")})
  
  return(non_trauma_pathway)
}
```

## 6. Modelling patient arrivals

Patients arrive a the urgent treatment centre following a time dependent process. When patients arrive they are classified as trauma or non-trauma.

To modify the classification of patients we will use a trajectory that uses the \`branch\` function from `simmer`.

The function \`sample_arrival_type\` returns a 1 (trauma) or 2 (non-trauma). This is used to select the appropriate patient trajectory.

```{r}

create_arrival_generator <- function(exp){
  
  DEPART_MSG <- "A patient has departed the UTC"
  
  # create and parameterise the trauma pathway trajectory
  trauma_pathway <- create_trauma_pathway(exp)
  
  # create and parameterise the non-trauma pathway trajectory
  non_trauma_pathway <- create_non_trauma_pathway(exp)
  
  patient_arrival <- trajectory() %>%
    branch(
      function() sample_arrival_type(exp$prob_trauma), continue=T,
        trauma_pathway,
        non_trauma_pathway
    ) %>%
    log_(function() {paste(DEPART_MSG)},level=1) %>% 
    set_attribute("departed", 1)
  
  return(patient_arrival)
}
```

## 7. Single run of the model

> Work in progress

```{r}
single_run <- function(env, exp, 
                       rep_number=1, 
                       run_length=DEFAULT_RESULTS_COLLECTION_PERIOD, 
                       debug_arrivals=FALSE){
  # add the simmer environment to the experiment list.
  exp <- c(exp, env=env) 
  
  # Create the arrivals generator
  arrival_gen <- create_arrival_generator(exp)
  
  # create model and run.
  env %>% 
    add_resource("triage_bay", exp$n_triage_bays) %>%
    add_resource("registration_clerk", exp$n_reg_clerks) %>%
    add_resource("examination_room", exp$n_exam_rooms) %>%
    add_resource("trauma_room", exp$n_trauma_rooms) %>%
    add_resource("trauma_treat_cubicle", exp$n_trauma_cubicles) %>%
    add_resource("nontrauma_treat_cubicle", exp$n_non_trauma_cubicles) %>%
    add_generator("Patient", arrival_gen, 
                  function() nspp_thinning(now(env), exp$arrival_data, 
                                           debug=debug_arrivals),
                  mon=2) %>% 
    run(until=run_length)
  
  # return environment and all of its results.
  return(env)
}
```

Script to conduct single run of the model

> Note that the environment is created outside of the `single_run` function. This is to separate the creation of the environment from the `run` function call. The reason is so that the `now(env)` function will work correctly in the `nspp_thinning` sampling function (if we do not separate then the same time is always passed to the function).

```{r}
set.seed(SEED)
exp <- create_experiment(log_level=0)
treat_sim <- simmer("TreatSim", log_level=exp$log_level)
treat_sim <- single_run(treat_sim, exp)
print("Simulation Complete.")
```

## 8. Multiple replications

```{r}
N_REPS <- 50

set.seed(42)

# note unlike in simmer documentation we use a traditional for loop
# instead of lapply. This allows us to separate env creation
# from run and preserve the environment interaction between NSPP 
# and current sim time.
# TO DO: look again -> can treat_sim be created inside single_run()
print("running replications...")
reps = vector()
for(rep in 1:N_REPS){
  exp <- create_experiment(log_level=0)
  treat_sim <- simmer("TreatSim", log_level=exp$log_level)
  treat_sim <- single_run(treat_sim, exp)
  # store the latest simulation enviornment and its results.
  reps <- c(reps, treat_sim)
}
print("Complete.")
```

## 9. Results analysis

Analysis of `simmer` results is achieved using a mix of statistics collected automatically and custom attributes set by the modeller during the run.

In general, we follow a typical strategy in a simulation study. We calculate the mean Key Performance Indicator (KPI) seen during an individual replication of the model (e.g. waiting time for triage and utilisation of the the triage rooms). This is repeated for all replications and the distribution of results can be visualised or we use a summary measure such as the mean.

Below we construct a summary table of results providing the mean of 16 KPIs.

> Code quality is a work in progress. 😃

```{r}
arrivals_by_replication <- function(envs){
  results <- vector()
  for(env in envs){
    results <- c(results, get_n_generated(env, "Patient"))
  }
  return(data.frame(results))
}
```

```{r}
get_resource_counts <- function(exp) {
  resource = c("triage_bay", 
                "registration_clerk", 
                "examination_room",
                "trauma_room",
                "trauma_treat_cubicle",
                "nontrauma_treat_cubicle")
  
  resource_counts = c(exp$n_triage_bays,
                      exp$n_reg_clerks,
                      exp$n_exam_rooms,
                      exp$n_trauma_rooms,
                      exp$n_trauma_cubicles,
                      exp$n_non_trauma_cubicles)
  
  df_resource <- data.frame(resource)
  df_resource$count <- resource_counts
  return(df_resource)
}
```

Functions to calculate KPIs from custom attributes

```{r}
# mean time in the system and throughput
calc_kpi_from_attributes_for_rep <- function(reps, rep){
  
  # get attributes
  att <- get_mon_attributes(reps)
  # for speed - limit to replication number.
  rep_att <- subset(att[att$replication == rep,], select = c(name, key, value))
  # long to wide format...
  data_wide <- spread(rep_att, key, value)
  
  # Patient type 1: trauma
  # take the mean and ignore patients still in pathway
  mean_time_1 = mean(data_wide[data_wide$patient_type == 1,]$total_time, na.rm = TRUE)
  
  # Patient type 2: non_trauma
  # take the mean and ignore patients still in pathway
  mean_time_2 = mean(data_wide[data_wide$patient_type == 2,]$total_time, na.rm = TRUE)
  
  # Throughput - discharges during opening hours.
  throughput <- sum(data_wide$departed, na.rm=TRUE)
  
  # store and return data.frame of results
  rep_results <- data.frame("05_total_time(non-trauma)" = mean_time_2,
                            "08_total_time(trauma)" = mean_time_1, 
                            "09_throughput"= throughput)
  return(rep_results)
}

calc_kpi_from_attributes <- function(reps){
  # calcs total time by patient type and total throughput
  
  # empty dataframe for attribute calculations.
  att_results <- data.frame(matrix(ncol = 3, nrow = 0))
  colnames(att_results) <- c("05_total_time(non-trauma)", "08_total_time(trauma)", "09_throughput")

  # add each rep separately as this works faster with pivot
  for(rep_i in 1:length(reps)){
    att_results <- rbind(att_results, calc_kpi_from_attributes_for_rep(reps, rep_i))
  }

  # return the KPIs by replications
  return(att_results)
}
```

```{r}
create_summary_table <- function(reps, exp, dp=2){

  # mean number of arrivals
  arrival_counts <- arrivals_by_replication(reps)
  mean_arrivals <- mean(arrival_counts$results)
  arrival_df <- data.frame("00_arrivals", mean_arrivals)
  names(arrival_df)<-c("kpi","mean")

  
  # get results dataframe broken down by resource and replication.
  arrivals <- get_mon_arrivals(reps, per_resource=TRUE)
  
  # calculate waiting time
  arrivals$waiting_time <-arrivals$end_time - arrivals$start_time - arrivals$activity_time
  # arrivals$waiting_time <- round(arrivals$waiting_time, 2)
  
  cols <- c("resource", "replication")
  summary_wait <- arrivals %>%
    # mean waiting time in each replication
    group_by(across(all_of(cols))) %>%
    summarise(rep_waiting_time=mean(waiting_time)) %>% 
    # mean waiting time across replications
    group_by(resource) %>% 
    summarise(mean=mean(rep_waiting_time)) %>% 
    mutate(resource=recode(resource,  
                         'triage_bay'='01a_triage_wait', 
                         'registration_clerk'='02a_registration_wait',
                         'examination_room'='03a_examination_wait',
                         'nontrauma_treat_cubicle'='04a_treatment_wait(non_trauma)',
                         'trauma_room'='06a_stabilisation_wait',
                         'trauma_treat_cubicle'='07a_treatment_wait(trauma)')) %>% 
    arrange(resource)
  
  
  names(summary_wait)<-c("kpi","mean")
  
  # utilisation calculation:
  # simple calculation of total busy time / total scheduled resource time.
  # where total scheduled time = n_resource * results collection period.
  summary_util <- arrivals %>%
    # mean waiting time in each replication
    group_by(across(all_of(cols))) %>%
    summarise(in_use=sum(activity_time)) %>% 
    arrange(resource)
  
  # add resource counts to table for calc
  summary_util <- merge(summary_util, get_resource_counts(exp), 
                        by="resource", all=TRUE)
  
  # total in use / total scheduled time
  summary_util$util = summary_util$in_use / (DEFAULT_RESULTS_COLLECTION_PERIOD * summary_util$count)  
  
  summary_util <- summary_util %>%
    # mean waiting time in each replication
    group_by(resource) %>%
    summarise(mean=mean(util)) %>% 
    mutate(resource=recode(resource,  
                         'triage_bay'='01b_triage_util', 
                         'registration_clerk'='02b_registration_util',
                         'examination_room'='03b_examination_util',
                         'nontrauma_treat_cubicle'='04b_treatment_util(non_trauma)',
                         'trauma_room'='06b_stabilisation_util',
                         'trauma_treat_cubicle'='07b_treatment_util(trauma)')) %>% 
    arrange(resource)

  names(summary_util) <- c("kpi","mean")
  
  overall <- bind_rows(arrival_df, summary_wait, summary_util) %>% 
    arrange(kpi) 
  
  ## KPIs calculated from attributes
  # empty dataframe for attribute calculations.
  att_results <- calc_kpi_from_attributes(reps)
  mean_att_results <- colMeans(att_results)

  overall <- overall %>%
    add_row(kpi="05_total_time(non-trauma)", mean=mean_att_results[1]) %>%
    add_row(kpi="08_total_time(trauma)", mean=mean_att_results[2]) %>%
    add_row(kpi="09_throughput", mean=mean_att_results[3]) %>%
    arrange(kpi)
  
  # round and return
  overall$mean <- round(overall$mean, dp)
  return(overall)
}
```

```{r}
results_summary <- create_summary_table(reps, exp)
results_summary
```
