# Sampling in R

By default R offers less control over random sampling than python. It uses a single random number stream for all sampling and does not allow you to create individual random number streams for each sampling distribution.

To illustrate the impact of this design consider the example below. Imagine we are going to simulate 5 patients length of stay (LoS) in an acute hospital followed by rehabilitation. Acute LoS is exponentially distribution while rehabilitation LoS follows a uniform distribution (the choice of distribution does not matter).

> IMPORTANT: To make the results "repeatable" we will set a random seed.

## Constants

```{r}
SEED <- 42
ACUTE_MEAN <- 32.0
REHAB_MIN <- 15.0
REHAM_MAX <- 80.0
```

## Experiment 1

```{r}
n_patients <- 5

set.seed(SEED)
acute_los <- rexp(n=n_patients, rate=1.0/ACUTE_MEAN)
rehab_los <- runif(n=n_patients, min=REHAB_MIN, max=REHAM_MAX)

print(acute_los)
print(rehab_los)
```

## Experiment 2

```{r}
n_patients <- 2

set.seed(SEED)
acute_los <- rexp(n=n_patients, rate=1.0/ACUTE_MEAN)
rehab_los <- runif(n=n_patients, min=REHAB_MIN, max=REHAM_MAX)

print(acute_los)
print(rehab_los)
```

```{r}
n_patients <- 5

set.seed(SEED)
acute_los <- rexp(n=1, rate=1.0/ACUTE_MEAN)
rehab_los <- runif(n=1, min=REHAB_MIN, max=REHAM_MAX)
```
