# Sampling in R

By default R offers less control over random sampling than python. It uses a single random number stream for all sampling and does not allow you to create individual random number streams for each sampling distribution.

To illustrate the impact of this design consider the example below. Imagine we are going to simulate 5 patients length of stay (LoS) in an acute hospital followed by rehabilitation. Acute LoS is exponentially distribution while rehabilitation LoS follows a uniform distribution (the choice of distribution does not matter).

> IMPORTANT: To make the results "repeatable" we will set a random seed.

## Constants

```{r}
SEED <- 42
ACUTE_MEAN <- 32.0
REHAB_MIN <- 15.0
REHAM_MAX <- 80.0
```

## Experiment 1

```{r}
n_patients <- 5

set.seed(SEED)
acute_los <- rexp(n=n_patients, rate=1.0/ACUTE_MEAN)
rehab_los <- runif(n=n_patients, min=REHAB_MIN, max=REHAM_MAX)

print(acute_los)
print(rehab_los)
```

## Experiment 2

We will now reset the random stream using the same seed and limit the number of patients simulated to 2.

When we re-run the code we might expect to get

> Acute Los:
>
> ```         
> 6.346778 21.148648
> ```
>
> Rehab Los:
>
> ```         
> 23.75333 57.70450
> ```

But we will see that this does not happen. This is because all sampling makes use of a pseudo random number stream that generates uniformly distribution numbers $U$'s between 0 and 1. When only 1 stream is used for all sampling we can end up with lots of noise between experiments simply because different $U$'s are used for the same patients.

```{r}
n_patients <- 2

set.seed(SEED)
acute_los <- rexp(n=n_patients, rate=1.0/ACUTE_MEAN)
rehab_los <- runif(n=n_patients, min=REHAB_MIN, max=REHAM_MAX)

print(acute_los)
print(rehab_los)
```

```{r}
n_patients <- 5

set.seed(SEED)

for (patient_i in 1:n_patients) {
    print(paste("Patient ", patient_i))
    print(paste("Acute ", rexp(n=1, rate=1.0/ACUTE_MEAN)))
    print(paste("Rehab ", runif(n=1, min=REHAB_MIN, max=REHAM_MAX)))
}


```

```{r}
n_patients <- 2

set.seed(SEED)

for (patient_i in 1:n_patients) {
    print(paste("Patient ", patient_i))
    print(paste("Acute ", rexp(n=1, rate=1.0/ACUTE_MEAN)))
    print(paste("Rehab ", runif(n=1, min=REHAB_MIN, max=REHAM_MAX)))
}

```
